Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-26 analysis/paper/paper.Rmd

Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced. Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local communities remains unanswered. Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017]. Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region (Andrade 2007; Kang 2012). Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period. Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century. Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.

We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan. Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network. We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan. We hypothesize that there is no change in the spatial distribution of ornaments at Kiwulan relating to these international contacts.

@@

Father Jacinto Esquivel recorded how people in Tamsui entered marriages. A man who wanted to take a wife had to give her parents cuentas. If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled (Li 2006: 151). Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals (Pao 2008: 122, 143, 151). Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity. They put a mat into the grave to protect the body from moisture and then put cooked rice on both sides of the head as provision. The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need. The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave (Li 2006: 153).

Documents from the Qing dynasty usually describe the decorative purposes of the ornaments. In the Kavalan Subprefecture Gazetteer (Chen 2006[1852]: 308), the Kavalan Zhi Lue (Ko 1993[1837]: 11, 126) and the Dong Cha Ji Lue (Yao 1996[1829]: 77), it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments. Their number was small and their value high, hence only the rich possessed them. Ordinary people wore agate beads or glass beads on their head or neck during various festivals. In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori (Kanori 1996: 227-232) made field surveys among various ethnic groups in Taiwan. He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.

Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans (Liu 2008: 133-134). Interview records describe agate beads as valuable objects passed down from mother to daughter. When they were not used, they were often put on roof beams so that they were not seen by other people. But the origin of these beads is still not known. Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer (Hsu 1992: 22) and carp-shaped ornaments had completely disappeared. Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination. It was usually the female shamans that were responsible for divination rites, a technique passed down by ancestors. Most aborigines today don’t know the origin of agate beads any more, but their worth is still highlighted by oral history and through their scarcity.

@@

Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period. The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution. We randomly assigned points, here the location of ornaments, for each square according to the data, subset three groups for three time periods, and compute the kernal density for each time period for comparison. Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space. Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier.etal]. Density value is assigned to each cell, of which the unit is meter. The results \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period. Those maps present three consistent sub-region with shifting core area over time. The distribution might indicate different social groups who own more ornaments, and the multiple groups during European contact period might reflect a more corporate mode of social organization. However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious core area. Next step is to test whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.

@@

@@

Q: what is the scale of those exchange network: 
- pre-European contact: middle scale, regional trade network 
- European contact: global scale, international trade network 
- Chinese contact: small scale, mutual trade network

Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
- pre-European contact: indirect 
- European contact: indirect, colonial context
- Chinese contact: direct, Han people migration

 and intense of contact

@@

>>>>

Compared to other places in Asia and Oceania, the impact of European colonialism on indigenous communities in East Asia appears to have been much less pronounced. Direct colonial rule was rare and limited, but the question of long-lasting indirect impacts on local communities remains unanswered. Indirect effects of colonialism is addressed recently to discuss the impact on the local indigenous societies in the periphery of colonial control [@Trabert2017]. Taiwan was colonized by the Spanish and the Dutch during the early 17th century, where northeastern Taiwan, Yilan, is an especially unique example for exploring indirect effects due to its physically isolated location that made it not easily accessible to the Europeans and the colonial control was less compared to other parts of this region [@Andrade2007; @Kang2012). Prior to the contact with the Europeans, there were small-scale regional exchange networks between China and Taiwan since in the Neolithic period. Also, northern Taiwan had been involved in long-distance networks of East Asia since 14th century, and encountered a large wave of Han Chinese migration in the 19th century. Northeastern Taiwan can provide a good example to explore the indirect impacts on indigenous societies caused by different foreign groups in different periods by taking a long-term perspective research.

We ask whether colonial influence on indigenous populations can be detected in ornaments recovered from the archaeological record at Kiwulan (1400-1900 AD), a large Iron Age settlement in northeastern Taiwan. Our hypothesis is that there was a greater diversity of ornaments types and materials at Kiwulan after European contact due to the large scale exchange network. We hypothesize that after Chinese contact there was a decline in the production, use and discard of ornaments at Kiwulan. We hypothesize that there are clustered distribution of ornaments at Kiwulan relating to these international contacts.

@@

Father Jacinto Esquivel recorded how people in Tamsui entered marriages. A man who wanted to take a wife had to give her parents cuentas. If a married man was found having cheated on his wife, he only had to pay the witness with agate or golden beads and the matter was settled [@Li2006:151]. Other records mention that the female shamans (majuorbol) in the tribe used agate beads for healing ceremonies in rituals [@Pao 2008: 122, 143, 151]. Regarding funerals, some indigenous people buried the dead under their houses or in the vicinity. They put a mat into the grave to protect the body from moisture and then put cooked rice on both sides of the head as provision. The deceased were buried in small graves with both knees bent, and on the grave they placed quivers with arrows, pottery, agate beads and other things that they would need. The greater the influence of the deceased, the more agate, pottery and cloths were placed on the grave [@Li 2006: 153].

Documents from the Qing dynasty usually describe the decorative purposes of the ornaments. In the Kavalan Subprefecture Gazetteer [@Chen2006[1852]: 308], the Kavalan Zhi Lue [@Ko1993[1837]: 11, 126) and the Dong Cha Ji Lue [@Yao1996[1829]: 77], it is mentioned that aborigines in Yilan used metal threads to weave so called golden carp-shaped ornaments. Their number was small and their value high, hence only the rich possessed them. Ordinary people wore agate beads or glass beads on their head or neck during various festivals. In 1895, at the beginning of Japanese rule over Taiwan, Ino Kanori [@Kanori1996: 227-232] made field surveys among various ethnic groups in Taiwan. He describes the Plains Aborigines from Yilan as not using carp-shaped ornaments any more, but mentions that older people still use beads ornaments.

Recent ethnographic research of the Kavalan tribe describes agate beads as used in divination ceremonies, the subli, by female shamans [@Liu2008: 133-134]. Interview records describe agate beads as valuable objects passed down from mother to daughter. When they were not used, they were often put on roof beams so that they were not seen by other people. But the origin of these beads is still not known. Later, because they were bought by Japanese and people in Taipei, agate beads became rarer and rarer [@Hsu1992: 22] and carp-shaped ornaments had completely disappeared. Modern ethnographic research shows that agate beads were not only used as decoration, but also played an important role in divination. It was usually the female shamans that were responsible for divination rites, a technique passed down by ancestors. Most aborigines today don’t know the origin of agate beads any more, but their worth is still highlighted by oral history and through their scarcity.

@@

## Spatial distribution
Point pattern analysis is a way to assess whether the distribution of ornaments presents some spatial pattern that can indicate the changes during culture contact period. The location where ornaments from were square-based recorded, thus we took the intensity approach that focuses on the average density of points across space to explore their distribution. We randomly assigned points, here the location of ornaments, for each square according to the data, subset three groups for three time periods, and compute the kernal density for each time period for comparison. Kernal density estimations (KDE) can estimate the probability of the density of events across space by creating a continuous, smooth density surface across space. Here we use KDE to visualize the event of the presence of ornaments that were represents by maps on which we are able to see core areas and surrounding neighborhoods [@Bonnier.etal]. Density value is assigned to each cell, of which the unit is meter. The results \@ref(fig:plot-all-densities) show that there is one major core area during the pre-European contact period, multiple core areas during European contact period, and a single core during the Chinese contact period. Those maps present three consistent sub-region with shifting core area over time. The distribution might indicate different social groups who own more ornaments, and the multiple groups during European contact period might reflect a more corporate mode of social organization. However, the generation of core areas might be biased due to small sample size, for example, a few ornaments found at one single square during the Chinese period could create obvious core area. Next step is to test whether the distribution of ornaments is randomly distributed to understand the underlying social indication related to ornaments.

@@

The introduction of ornaments to Northeastern Taiwan during different periods could inform how local indigenous people perceive and distribute the trading goods in different contexts of culture contacts by examining their frequency and spatial distribution across space. Taiwan island had been involved in regional exchange networks of Southeast Asia since Neolithic period, a global trade network during the European contact period in the early 17th century, and later intense but smaller scale network with Han Chinese communities since 19th century [@Chen200?]. Kiwulan demonstrates continuous occupation from the late Iron Age to one hundred years ago over 600 years that covers these three different culture contacts that enables a comparative research to explore the indirect influence of culture contact especially in the global colonial context. These three analytic chronology at Kiwulan represent different scale of exchange network and intensity of culture contact. For the scale of trade network, Pre-European contact can be defined as a middle scale represented by the prehistoric regional trade network in Southeast Asia. International trade network during the European contact presents a global scale, while local trade network during the Chinese contact presents a small scale. For the intensity of culture contact, both pre-European contact and European contact period could be defined as indirect contact, but the colonial power underlying the contact with the Europeans should be also considered. The culture contact during the Chinese period is more direct due to the migration of Han Chinese to Kiwulan.

The results show there was a greater diversity of ornaments types and materials at Kiwulan during European contact. This might result from the large scale exchange network that stimulate the circulation of different ornaments classes. Among those ornaments, agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads, which might indicate the diverse origins. However, the frequency of overall ornaments and each subtype declines significantly during the Chinese contact period. The reason could be related to smaller scale of network and the overall decline of Kiwulan population at that time. The spatial pattern of ornaments shows that there were some clusters appeared since pre-European and European contact period. The hypothesis testing for spatial pattern further indicate more concentration of ornaments during European contact. Because trade ornaments had been viewed as prestige goods in local cultural context in the prehistoric Northeastern Taiwan, this differentiation could reflect some level of social inequality appeared before European contact and then it was reinforced during European period.

Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.

@@

@@

Indirect influence of colonialism has been proposed recently and some students in many parts of the world [@Trabert2017]. Kiwulan in Northeastern Taiwan is an especially important case study where the location was relatively isolated and peripheral that can provide valuable insight into the discussion of influence of indirect colonial contacts.
